### 2023-06-30 paper/main.tex
corrections

\subsection{Conventional 2.5D Wave Field Synthesis}
Consider a smooth convex secondary source distribution located at $\vxo = \posvec{3}{x_0}{y_0}{0}$ consisting of a continuous distribution of 3D point sources, described by the 3D freefield Green's function, reading in the frequency domain as
\begin{equation}
    G_{3 \mathrm{D}}(\vx,\vxo,\omega) = \frac{1}{4\pi}\frac{\te^{-\frac{\omega}{c}\left| \vx-\vxo\right|}}{\left| \vx-\vxo\right|},
\end{equation}
where $\omega$ denotes the angular frequency.
In this geometry the target field inside the area bounded by the SSD is described by the  Kirchhoff approximation of 2.5D Kirchhoff-Helmholtz integral, from which the 2.5D driving functions can be extracted.
For an arbitrary simple target virtual sound field $P(\vxo,\omega)$ the steady-state driving functions read as \cite{Firtha2016}
\begin{equation}
    D(\vxo,\omega) = \underbrace{\sqrt{8\pi \frac{\ti \omega}{c}}}_{ H_{\text{pre}}(\omega) }
    w(\vxo) \sqrt{d_{\mathrm{ref}}(\vxo)}
    P(\vxo,\omega)
    .
    \label{Eq:WFS25D}
\end{equation}
The driving function consists of
\begin{itemize}
    \item a frequency dependent pre-equalization filter $H_{\text{pre}}(\omega)$ \cite{Start1997:phd, doi:10.1121/1.404755, Bleistein1986},
    \item a secondary source selection window \cite{Nicol1999, Spors2007}
          \begin{equation}
              w(\vxo) = \max{\left( \hat{k}_{\mathrm{n}}^P(\vxo),0\right)},
          \end{equation}
          where $\hat{k}_{\mathrm{n}}^P(\vxo)$ is the normal component of the local propagation vector,
    \item a gain factor $\sqrt{d_{\mathrm{ref}}(\vxo)}$ allowing amplitude correction along a reference curve, depending on the actual virtual field model (c.f. \cite{Start1997:phd} and for a more complete treatise \cite{Firtha2016}),
    \item and the virtual field measured on the SSD.
\end{itemize}
Assuming a simple virtual sound field as given by \eqref{eq:Pxw} the driving functions can be written as
\begin{equation}
    D(\vxo,\omega) = H_{\text{pre}}(\omega)
    \underbrace{ w(\vxo)  \, \sqrt{d_{\mathrm{ref}}(\vxo)} \, A(\vxo) }_{A^D(\vxo)} \te^{- \ti \omega \tau(\vxo)}
    ,
\end{equation}
and in the time domain as
\begin{equation}
    d(\vxo,t) = h_{\text{pre}}(t) \ast_t  A^D(\vxo)\,  \delta\left( t - \tau(\vxo)\right),
    \label{eq:d_wfs_td}
\end{equation}
with $A^D(\vxo)$ being the real valued overall gain factor of the driving function, $\ast_t$ denoting temporal convolution and $h_{\text{pre}}(t)$ is the temporal WFS prefilter impulse response.
The analytical, continuous pre-equalization filter impulse response is given in \cite{Deregowski1983, Firtha2019phd}, while for an ideal, non-casual FIR filter implementation the reader is referred to \cite{Schultz2016}.
In the following due to the associativity of convolution this prefiltering is excluded from the discussion.

@@

Equation \eqref{eq:temporal_transform} can be directly expressed in the spectral domain by taking the Fourier transform of both sides, relating the frequency response of the spatial and the temporal filters.
By denoting $\mathcal{F}\left( h_x(s)\right) = H_x(k_s)$ the corresponding transform is given by
\begin{equation}
    H_t\left(s,\omega\right) = H_x\left(-\frac{\omega}{c}\hat{k}_t^P(s)\right) = H_x\left(-k_t^P(s)\right),
    \label{eq:frequency_transform}
\end{equation}
where $k_t^P(s)$ is the tangential component of the local wavenumber vector, being a vector in a steady state sound field, pointing in the local propagation direction with the length being $\omega/c$ \cite{Firtha2019phd}.
Hence, as the main result of the present discussion, the temporal filter transfer can be obtained from the wavenumber (modal) content of the spatial filter by simple rescaling in terms of the local wavenumber vector.

The validity and the error analysis of the above filter transform was investigated in details in \cite{Firtha DAGA2023}.
In the following the direct application for local WFS is discussed.

@@

Figure \ref{Fig:aliasing} (d) reflects that the overlapping of the sampled driving function spectra, and, therefore, the presence of aliasing waves can be avoided by spatially bandlimiting the driving functions to the Nyquist wavenumber prior to sampling.
Spatial bandlimitation can be straightforwardly performed by designing a suitable spatial antialiasing filter and transforming it into an equivalent temporal filter bank, based on the results of the previous section.
An important advantage of the above temporal filtering is that the driving functions may be evaluated in the discrete SSD positions, and a subsequent temporal antialiasing inherently ensures analytical spatial bandlimitation\footnote{
    The above statement holds as long as temporally continuous driving functions are assumed. Once the driving functions are sampled with a sampling frequency $f_s = 1 /\Delta t$, this corresponds to a spatial sampling prior to the filtering process with $\Delta s = c/f_s$.}.

@@

The result of antialiasing filtering is depicted in Figure \ref{Fig:symm_antialiasing} for the synthesis scenario discussed in the foregoing.
In the present case the sampling arc length is given by $\Delta s = \frac{2 \pi R_0}{N}$, i.e. the space-dependent cut-off frequency reads as 
\begin{equation}
    \omega_c(s) = \frac{ N}{ 2 R_0 } \frac{c}{\hat{k}_t^P(s)}.
\end{equation}
The antialiasing filters were a 4-th order Butterworth deisgn.

Figure (a) shows the result of the synthesis, while Figure (b) illustrates the spectrum of the bandlimited, sampled driving functions.
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width = 0.75\columnwidth]{figs/firtha_center.png}
            \footnotesize \put(0,2){(a)}
        \end{overpic}
        \\
        \begin{overpic}[width = 0.75\columnwidth]{figs/antialiased_synthesis_spectrum.png}
            \footnotesize \put(0,0){(b)}
        \end{overpic}
    \end{center}
    \caption{Symmetric antialiasing filtering. Figure (a) depicts the synthesized wave front, and (b) illustrates the spatial spectrum of the driving functions.}
    \label{Fig:symm_antialiasing}
\end{figure}
It is highlighted that the presented antialiasing strategy ensures suppressed aliasing waves in the center of the circular source array.
In other positions of the listening area lateral aliasing wavefronts remain in the reproduced wave field, while the target, virtual wave front is bandlimited.
This phenomena can be easily understood by investigating the synthesis scenario in a geometrical manner:

Since the cutoff frequency of the antialiasing filter (Eq. \eqref{eq:cutoff_fr} ) is inversely proportional to the tangential component of the virtual field on the SSD, therefore, in a circular array a single SSD element will have nearly fullband driving signal: that SSD element at which the wavefront arrives perpendicularly.
This fullband SSD element is marked with white dot in Figure \ref{Fig:symm_antialiasing} (a).
From general WFS theory it is known that the fullband SSD element will dominate the synthesized wavefront (i.e. will serve as a stationary SSD position) for spatial positions along a straight line crossing the actual SSD element into the direction of the local propagation vector.
Therefore, the present antialiasing strategy ensures aliasing-free synthesis along a straight line for which the fullband SSD element is the stationary secondary source.
The direction of the line of aliasing-free synthesis is marked with a dashed arrow in the figure.
This is in contrast with previous local WFS approaches, in which the synthesis is optimized in the proximity of a pre-defined reference position over a circular area.

It should be noted that in the angular frequency domain the filtered driving function is written as
\begin{equation}
    D'_t(s,\omega) = D(s,\omega) H_t(s,\omega) =  H_x\left(-k_t^P(s)\right) \, D(s,\omega),
\end{equation}
from which formulation it is apparent that in the frequency domain the presented spatial-to-temporal filter transform strategy can be also interpreted as a frequency dependent spatial windowing approach, and the frequency dependent spatial window is given by $H_x\left(-k_t^P(s)\right)$:
in the aspect of antialiasing filtering SSD positions where the local wavenumber vector is higher than the Nyquist wavenumber are highly attenuated, with the window width decreasing with increasing angular frequency.
\begin{figure}[]
    \begin{overpic}[width = 1\columnwidth]{figs/equivalent_windows.png}
    \end{overpic}
    \caption{Illustration of equivalent frequency dependent window functions}
    \label{Fig:equivalent_windows}
\end{figure}
The actual shape of the spatial windows are given by the wavenumber spectrum of the spatial lowpass filter $H_x(k_s)$, rescaled in terms of the local wavenumber vector.
The set of window functions applied in the present simulation scenario are illustrated in Figure \ref{Fig:equivalent_windows}.

>>>>

\subsection{Conventional 2.5D Wave Field Synthesis}
Consider a smooth convex secondary source distribution located at $\vxo = \posvec{3}{x_0}{y_0}{0}$ consisting of a continuous distribution of 3D point sources, described by the 3D freefield Green's function, reading in the frequency domain as
\begin{equation}
    G_{3 \mathrm{D}}(\vx,\vxo,\omega) = \frac{1}{4\pi}\frac{\te^{-\frac{\omega}{c}\left| \vx-\vxo\right|}}{\left| \vx-\vxo\right|},
\end{equation}
where $\omega$ denotes the angular frequency.
In this geometry the target field inside the area bounded by the SSD is described by the  Kirchhoff approximation of 2.5D Kirchhoff-Helmholtz integral, from which the 2.5D driving functions can be extracted.
For an arbitrary simple target virtual sound field $P(\vxo,\omega)$ the steady-state driving functions read as \cite{Firtha2016}
\begin{equation}
    D(\vxo,\omega) = \underbrace{\sqrt{8\pi \frac{\ti \omega}{c}}}_{ H_{\text{pre}}(\omega) }
    w(\vxo) \sqrt{d_{\mathrm{ref}}(\vxo)}
    P(\vxo,\omega)
    .
    \label{Eq:WFS25D}
\end{equation}
The driving function consists of
\begin{itemize}
    \item a frequency dependent pre-equalization filter $H_{\text{pre}}(\omega)$ \cite{Start1997:phd, Berkhout1993:Acoustic_control_by_WFS, Bleistein1986},
    \item a secondary source selection window \cite{Nicol1999, Spors2007}
          \begin{equation}
              w(\vxo) = \max{\left( \hat{k}_{\mathrm{n}}^P(\vxo),0\right)},
          \end{equation}
          where $\hat{k}_{\mathrm{n}}^P(\vxo)$ is the normal component of the local propagation vector,
    \item a gain factor $\sqrt{d_{\mathrm{ref}}(\vxo)}$ allowing amplitude correction along a reference curve, depending on the actual virtual field model (c.f. \cite{Start1997:phd} and for a more complete treatise \cite{Firtha2016}),
    \item and the virtual field measured on the SSD.
\end{itemize}
Assuming a simple virtual sound field as given by \eqref{eq:Pxw} the driving functions can be written as
\begin{equation}
    D(\vxo,\omega) = H_{\text{pre}}(\omega)
    \underbrace{ w(\vxo)  \, \sqrt{d_{\mathrm{ref}}(\vxo)} \, A(\vxo) }_{A^D(\vxo)} \te^{- \ti \omega \tau(\vxo)}
    ,
\end{equation}
and in the time domain as
\begin{equation}
    d(\vxo,t) = h_{\text{pre}}(t) \ast_t  A^D(\vxo)\,  \delta\left( t - \tau(\vxo)\right),
    \label{eq:d_wfs_td}
\end{equation}
with $A^D(\vxo)$ being the real valued overall gain factor of the driving function, $\ast_t$ denoting temporal convolution and $h_{\text{pre}}(t)$ is the temporal WFS prefilter impulse response.
The analytical, continuous pre-equalization filter impulse response is given in \cite{Deregowski1983, Firtha2019phd}, while for an ideal, non-casual FIR filter implementation the reader is referred to \cite{Schultz2016}.
In the following due to the associativity of convolution this prefiltering is excluded from the discussion.

@@

Equation \eqref{eq:temporal_transform} can be directly expressed in the spectral domain by taking the Fourier transform of both sides, relating the frequency response of the spatial and the temporal filters.
By denoting $\mathcal{F}\left( h_x(s)\right) = H_x(k_s)$ the corresponding transform is given by
\begin{equation}
    H_t\left(s,\omega\right) = H_x\left(-\frac{\omega}{c}\hat{k}_t^P(s)\right) = H_x\left(-k_t^P(s)\right),
    \label{eq:frequency_transform}
\end{equation}
where $k_t^P(s)$ is the tangential component of the local wavenumber vector \footnote{The local wavenumber vector can be defined for a steady state sound field, pointing in the local propagation direction with the length being $\omega/c$ \cite{Firtha2019phd}}.
Hence, as the main result of the present discussion, the temporal filter transfer can be obtained from the wavenumber (modal) content of the spatial filter by simple rescaling in terms of the local wavenumber vector.

The validity and the error analysis of the above filter transform was investigated in details in \cite{firtha_daga2023}.
In the following the direct application for local WFS is discussed.

@@

Figure \ref{Fig:aliasing} (d) reflects that the overlapping of the sampled driving function spectra, and, therefore, the presence of aliasing waves can be avoided by spatially bandlimiting the driving functions to the Nyquist wavenumber prior to sampling.
Spatial bandlimitation can be straightforwardly performed by designing a suitable spatial antialiasing filter and transforming it into an equivalent temporal filter bank, based on the results of the previous section.
An important advantage of the temporal filtering approach instead of direct spatial filtering is that analyical spatial bandlimitation can be achieved by calculating the WFS driving functions at discrete positions, followed by temporal filtering \footnote{
The above statement holds as long as temporally continuous driving functions are assumed. Once the driving functions are sampled with a sampling frequency $f_s = 1 /\Delta t$, this corresponds to a spatial sampling prior to the filtering process with $\Delta s = c/f_s$.}.

@@

The result of antialiasing filtering is depicted in Figure \ref{Fig:symm_antialiasing} for the synthesis scenario discussed in the foregoing.
In the present case the sampling arc length is given by $\Delta s = \frac{2 \pi R_0}{N}$, i.e. the space-dependent cut-off frequency reads as 
\begin{equation}
    \omega_c(s) = \frac{ N}{ 2 R_0 } \frac{c}{\hat{k}_t^P(s)}.
\end{equation}
The antialiasing filters are 4-th order Butterworth designs.

Figure (a) shows the result of the synthesis, while Figure (b) illustrates the spectrum of the bandlimited, sampled driving functions.
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width = 0.75\columnwidth]{figs/firtha_center.png}
            \footnotesize \put(0,2){(a)}
        \end{overpic}
        \\
        \begin{overpic}[width = 0.75\columnwidth]{figs/antialiased_synthesis_spectrum.png}
            \footnotesize \put(0,0){(b)}
        \end{overpic}
    \end{center}
    \caption{Symmetric antialiasing filtering. Figure (a) depicts the synthesized wave front, and (b) illustrates the spatial spectrum of the driving functions.}
    \label{Fig:symm_antialiasing}
\end{figure}
It is highlighted that the presented antialiasing strategy ensures suppressed aliasing waves in the center of the circular source array.
In other positions of the listening area lateral aliasing wavefronts remain in the reproduced wave field, while the target, virtual wave front is bandlimited.
This phenomenon can be easily understood by investigating the synthesis scenario in a geometrical manner:

Since the cutoff frequency of the antialiasing filter (Eq. \eqref{eq:cutoff_fr}) is inversely proportional to the tangential component of the virtual field on the SSD, therefore, in a circular array a single SSD element will have nearly fullband driving signal: that SSD element at which the wavefront arrives perpendicularly (i.e. where the tangential component of the local propagation vector is zero).
This fullband SSD element is marked with white dot in Figure \ref{Fig:symm_antialiasing} (a).
From general WFS theory it is known that the fullband SSD element will dominate the synthesized wavefront (i.e. will serve as a stationary SSD position) for spatial positions along a straight line crossing the actual SSD element into the direction of the local propagation vector.
Therefore, the present antialiasing strategy ensures aliasing-free synthesis along a straight line for which the fullband SSD element is the stationary secondary source.
The direction of the line of aliasing-free synthesis is marked with a dashed arrow in the figure.
This is in contrast with previous local WFS approaches, in which the synthesis is optimized in the proximity of a pre-defined reference position over a circular area.

It should be noted that in the angular, temporal frequency domain the filtered driving function is written as
\begin{equation}
    D'_t(s,\omega) = D(s,\omega) H_t(s,\omega) =  H_x\left(-k_t^P(s)\right) \, D(s,\omega),
\end{equation}
from which formulation it is apparent that in the frequency domain the presented spatial-to-temporal filter transform strategy can be also interpreted as a frequency dependent spatial windowing approach, and the frequency dependent spatial window is given by $H_x\left(-k_t^P(s)\right)$:
in the aspect of antialiasing filtering SSD positions where the local wavenumber vector is higher than the Nyquist wavenumber are highly attenuated, with the window width decreasing with increasing angular frequency.
\begin{figure}[]
    \begin{overpic}[width = 1\columnwidth]{figs/equivalent_windows.png}
    \end{overpic}
    \caption{Illustration of equivalent frequency dependent window functions}
    \label{Fig:equivalent_windows}
\end{figure}
The actual shape of the spatial windows are given by the wavenumber spectrum of the spatial lowpass filter $H_x(k_s)$, rescaled in terms of the local wavenumber vector.
The set of window functions applied in the present simulation scenario are illustrated in Figure \ref{Fig:equivalent_windows}.
